Traffic Collisions in Toronto: Analysis of Key Variables

### 2024-09-26 paper/paper.qmd
update paper

---
title: "Traffic Collisions in Toronto: Analysis of Key Variables"
subtitle: "An Analysis of Traffic Collision Data from 2014-2024"
author: 
  - Andy Jiang
thanks: "Code and data are available at: https://github.com/AndyYanxunJiang/toronto-traffic-collisions.git"
date: 24 September 2024
date-format: long
abstract: "This paper analyzes traffic collision data in Toronto from 2014 to 2024. It explores trends in collisions by month, year, and neighborhood, examining variables such as fatalities, injuries, and property damage. The analysis is done using R, with visualizations produced through ggplot2."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(knitr)

@@

# Optional: Convert relevant variables to factors
data <- data %>%
  mutate(OCC_MONTH = factor(OCC_MONTH, levels = month.name),
         INJURY_COLLISIONS = as.factor(INJURY_COLLISIONS),
         FTR_COLLISIONS = as.factor(FTR_COLLISIONS),
         PD_COLLISIONS = as.factor(PD_COLLISIONS),
         AUTOMOBILE = as.factor(AUTOMOBILE),
         PEDESTRIAN = as.factor(PEDESTRIAN))
```

@@

head(data, 10) |>
  kable(
    col.names = c("Month", "Day of Week", "Year", "Hour", "Fatalities",
                  "Injury Collisions", "FTR Collisions", "PD Collisions", 
                  "Neighbourhood", "Automobile", "Motorcycle", 
                  "Passenger", "Bicycle", "Pedestrian"),
    booktabs = TRUE
  )
```

@@

```{r}
#| label: fig-collisions-month
#| fig-cap: Collisions by Month
#| echo: false

data %>% 
  ggplot(aes(x = OCC_MONTH, fill = factor(OCC_YEAR))) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Traffic Collisions by Month (2014-2024)",
       x = "Month", y = "Count", fill = "Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Collisions by Hour of the Day

```{r}
#| label: fig-collisions-hour
#| fig-cap: Collisions by Hour of the Day
#| echo: false

data %>% 
  ggplot(aes(x = OCC_HOUR, fill = factor(OCC_YEAR))) +
  geom_histogram(binwidth = 1, position = "dodge") +
  theme_minimal() +
  labs(title = "Traffic Collisions by Hour of the Day (2014-2024)",
       x = "Hour", y = "Count", fill = "Year")

@@

```{r}
#| label: fig-monthly-collisions
#| fig-cap: Monthly distribution of traffic collisions
#| echo: false

data %>%
  count(OCC_MONTH) %>%
  ggplot(aes(x = OCC_MONTH, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Monthly Distribution of Traffic Collisions", 
       x = "Month", y = "Number of Collisions")

@@

data %>%
  group_by(OCC_YEAR) %>%
  summarize(Total_Fatalities = sum(FATALITIES, na.rm = TRUE)) %>%
  ggplot(aes(x = OCC_YEAR, y = Total_Fatalities)) +
  geom_line(color = "darkred", size = 1) +
  theme_minimal() +
  labs(title = "Fatalities by Year", x = "Year", y = "Number of Fatalities")

>>>>

---
title: "Traffic Collisions in Toronto: Analysis of Key Variables"
subtitle: "An Analysis of Traffic Collision Data from 2014-2024"
author: 
  - Andy Jiang
thanks: "Code and data are available at: https://github.com/AndyYanxunJiang/toronto-traffic-collisions.git"
date: 24 September 2024
date-format: long
abstract: "This paper analyzes traffic collision data in Toronto from 2014 to 2024. It explores trends in collisions by month, year, and neighborhood, examining variables such as fatalities, injuries, and property damage. The analysis is done using @citeR, with visualizations produced through ggplot2."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(knitr)
library(kableExtra)
library(opendatatoronto)
library(here)
library(tidyr)
library(dplyr)
library(janitor)
library(VennDiagram)

@@

# Convert relevant variables to factors
data <- data %>%
  mutate(OCC_MONTH = factor(OCC_MONTH, levels = month.name),
         INJURY_COLLISIONS = as.factor(INJURY_COLLISIONS),
         FTR_COLLISIONS = as.factor(FTR_COLLISIONS),
         PD_COLLISIONS = as.factor(PD_COLLISIONS))
```

@@

# Display the cleaned collision data
head(data, 10) %>%

  # Use kable to format the table with custom column names
  kable(
    col.names = c(
      "Month", "Day of Week", "Year", "Hour", "Fatalities",
      "Injury Collisions", "FTR Collisions", "PD Collisions"
    ),
    align = 'c'
  )
```

@@

```{r}
#| label: fig-collisions-month
#| fig-cap: Monthly traffic collision counts in Toronto from 2014 to 2024.
#| echo: false
ggplot(data, aes(x = OCC_MONTH)) +
  geom_bar(position = "dodge", stat = "count", fill = "lightblue") +
  theme_minimal() +
  labs(title = "Total Traffic Collisions by Month (2014-2024)",
       x = "Month", y = "Total Count of Collisions") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, color = "black"),
        plot.title = element_text(hjust = 0.5, size = 14))
```

```{r}
#| label: fig-collisions-hour
#| fig-cap: Total collisions by hour of the day from 2014 to 2024
#| echo: false

data$OCC_HOUR[is.na(data$OCC_HOUR)] <- 0
data %>%
  group_by(OCC_HOUR) %>%
  summarise(Total_Collisions = sum(INJURY_COLLISIONS == "YES" | 
                                     FTR_COLLISIONS == "YES" | 
                                     PD_COLLISIONS == "YES", na.rm = TRUE)) %>%
  ggplot(aes(x = OCC_HOUR, y = Total_Collisions)) +
  geom_line(color = "steelblue", size = 1) +
  theme_minimal() +
  scale_x_continuous(breaks = 0:23) +
  labs(title = "Total Traffic Collisions by Hour of the Day (2014-2024)",
       x = "Hour of the Day",
       y = "Total Number of Collisions") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        panel.grid.major.x = element_line(size = 1.2),
        panel.grid.major.y = element_line(size = 0.5))
```

```{r}

@@

```{r}
#| label: fig-collisions-type
#| fig-cap: Proportion of each type of traffic collisions (2014–2024).
#| echo: false

collision_summary <- data %>%
  summarise(Injury_Collisions = sum(INJURY_COLLISIONS == "YES"),
            FTR_Collisions = sum(FTR_COLLISIONS == "YES"),
            PD_Collisions = sum(PD_COLLISIONS == "YES"))

collision_summary_long <- collision_summary %>%
  pivot_longer(cols = everything(), names_to = "Collision_Type", 
               values_to = "Count") %>%
  mutate(Collision_Type = recode(Collision_Type, 
                                 "Injury_Collisions" = "Injury Collisions",
                                 "FTR_Collisions" = 
                                   "Failed to Remain Collisions",
                                 "PD_Collisions" = 
                                   "Property Damage Collisions"))

collision_summary_long <- collision_summary_long %>%
  mutate(Percentage = round(Count / sum(Count) * 100, 2))

ggplot(collision_summary_long, aes(x = "", y = Count, fill = Collision_Type)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(Percentage, "%")), 
            position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("Injury Collisions" = "lightblue", 
                               "Failed to Remain Collisions" = "skyblue", 
                               "Property Damage Collisions" = "#4682B4")) +
  theme_void() +
  labs(title = "Proportion of Each Collision Type") +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))
```

data %>%
  group_by(OCC_YEAR) %>%
  summarize(Total_Fatalities = sum(FATALITIES, na.rm = TRUE)) %>%
  ggplot(aes(x = OCC_YEAR, y = Total_Fatalities)) +
  geom_line(color = "darkred", size = 1) +
  scale_x_continuous(breaks = seq(min(data$OCC_YEAR), 
                                  max(data$OCC_YEAR), by = 1)) +  
  theme_minimal() +
  labs(title = "Fatalities by Year", 
       x = "Year", 
       y = "Number of Fatalities") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 14),
    panel.grid.major.x = element_line(size = 0.8, color = "lightgrey")
  )
```

```{r}

@@

```{r}

```{r}
